---
title: "Running CAPM for all DJIA components"
author: "Kostis Christodoulou"
date: "Nov 2021"
output:
  html_document:
    theme: flatly
    highlight: zenburn
    toc: yes
    toc_float: yes
---


```{r huxtable-stuff, include=FALSE}
options("huxtable.knit_print_df" = FALSE)
options(digits = 3)
library(janitor)
library(lubridate)
library(here)
library(tidyverse)
library(ggfortify)
library(mosaic)
library(huxtable)
library(kableExtra)
library(tidyquant)
library(rvest)
library(ggridges)
library(broom)
library(forcats)
```

# Using `purr::map()` to estimate all betas of the 30 DJIA stocks 

We will get a list for all of the [Dow Jones Industrial Average (DJIA)](https://en.wikipedia.org/wiki/Dow_Jones_Industrial_Average) stocks and their ticker symbols and download some data. Besides the thirty stocks that make up the DJIA, we will also add `SPY` which is an SP500 ETF (Exchange Traded Fund).

## Get components off wikipedia 

```{r, tickers_from_wikipedia}

djia_url <- "https://en.wikipedia.org/wiki/Dow_Jones_Industrial_Average"

#get tables that exist on URL
tables <- djia_url %>% 
  read_html() %>% 
  html_nodes(css="table")


# parse HTML tables into a dataframe called djia. 
# Use purr::map() to create a list of all tables in URL
djia <- map(tables, . %>% 
               html_table(fill=TRUE)%>% 
               clean_names())


# constituents
table1 <- djia[[2]] %>% # the second table on the page contains the ticker symbols
  mutate(date_added = ymd(date_added),
         
         # if a stock is listed on NYSE, its symbol is, e.g., NYSE: MMM
         # We will get prices from yahoo finance which requires just the ticker
         
         # if symbol contains "NYSE*", the * being a wildcard
         # then we jsut drop the first 6 characters in that string
         ticker = ifelse(str_detect(symbol, "NYSE*"),
                          str_sub(symbol,7,11),
                          symbol)
         )

# we need a vector of strings with just the 30 tickers + SPY
tickers <- table1 %>% 
  select(ticker) %>% 
  pull()  # pull() gets them as a sting of characters


```


## Get price data using `tidyquant`


```{r get_price_data, message=FALSE, warning=FALSE, cache=TRUE}
# Notice the cache=TRUE argument in the chunk options. Because getting data is time consuming, # cache=TRUE means that once it downloads data, the chunk will not run again next time you knit your Rmd

starting_date <- Sys.Date() - 365*5 # go back 5 years from today; otherwise, give starting date as "2017-01-01"
ending_date <- Sys.Date()

myStocks <- tickers %>% 
  tq_get(get  = "stock.prices",
         from = starting_date, # go back 5 years from today; otherwise, give starting date as "2017-01-01"
         to = ending_date
         ) %>%
  group_by(symbol)


# get prices for SPY, the SP500 ETF
spy <- tq_get("SPY", get  = "stock.prices",
         from = starting_date, # go back 5 years from today; otherwise, give starting date as "2017-01-01"
         to = ending_date ) 

glimpse(myStocks) # examine the structure of the resulting data frame
```

For each ticker symbol, the data frame contains its `symbol`, the `date`, the prices for `open`,`high`, `low` and `close`, and the `volume`, or how many stocks were traded on that day. More importantly, the data frame contains the `adjusted` closing price, which adjusts for any stock splits and/or dividends paid and this is what we will be using for our analyses.

## Calculate monthly returns

Financial performance and CAPM analysis depend on __returns__ and not on __adjusted closing prices__. If I buy a stock today for 100 and I sell it tomorrow for 101.75, my one-day return, assuming no transaction costs, is 1.75%. So given the adjusted closing prices, our first step is to calculate daily and monthly returns.

```{r calculate_returns, message=FALSE, warning=FALSE, cache=TRUE}

#calculate monthly  returns for the 30 DJIA stocks
myStocks_returns_monthly <- myStocks %>% 
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "monthly_return",
               cols = c(nested.col)) 

#calculate SPY monthly  returns
spy_returns_monthly <- spy %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "SPY_return",
               cols = c(nested.col))

DJIA_data <- left_join(myStocks_returns_monthly, spy_returns_monthly, by="date")

```

## Use `purrr:map()` to apply an `lm()` function to a list of stocks

```{r djia_CAPM, warning=FALSE}
many_models <- DJIA_data %>% 
  # create a dataframe containing a separate dataframe for each stock
  group_by(symbol) %>% 
  nest()  %>% 
  
  # Run a simple regression model for every stock in the dataframe
  mutate(simple_capm_model = data %>% 
           map(~lm(monthly_return ~  SPY_return, data = .))) %>% 
  
  # extract coefficients and model details with broom::tidy
  mutate(coefs = simple_capm_model %>% 
           map(~ tidy(., conf.int = TRUE)),
         details = simple_capm_model %>% map(glance)) %>% 
  ungroup()

# pull intercepts, or alphas 
intercepts <- 
  many_models %>% 
  unnest(coefs) %>% 
  filter(term == "(Intercept)") %>% 
  arrange(estimate) %>% 
  mutate(symbol = fct_inorder(symbol)) %>% 
  select(symbol, estimate, std.error, conf.low, conf.high)

# let us look at the intercepts, or alphas
intercepts %>% 
  rename(alpha = estimate) %>% 
  kable()%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


# pull slopes, or betas 
slopes <- many_models %>% 
  unnest(coefs) %>% 
  filter(term == "SPY_return") %>% 
  arrange(estimate) %>% 
  mutate(symbol = fct_inorder(symbol)) %>% 
  select(symbol, estimate, std.error, conf.low, conf.high)

# let us look at the betas, or measure of market risk
slopes %>% 
  rename(beta = estimate) %>% 
  arrange(desc(beta)) %>% 
  kable()%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

# plot betas
ggplot(data = slopes, aes(x = symbol, y = estimate))+
  geom_col()+
  coord_flip()+
  theme_minimal(9)+
    labs(title = 'Estimating betas of stocks in the DJIA',
       subtitle = paste0(zoo::as.yearmon(starting_date), " to ", zoo::as.yearmon(ending_date)),
       x = "",
       y = "beta estimate")+
  
  geom_hline(yintercept = 1)+
  geom_label(aes(x = symbol, y = estimate, label = round(estimate, 2)),
             hjust = 1, 
             vjust = 0.5, 
             colour = "white", 
             fill = NA, 
             label.size = NA, 
             family="Helvetica", 
             size = 3)+
  NULL



# pull R^2 and regression SE, or specific risk 
rsq_errors <- many_models %>% 
  unnest(details) %>% 
  select(symbol, r.squared, sigma) %>% 
  arrange(desc(r.squared)) 

# let us join everything together
capm_results <- inner_join(slopes, intercepts, by="symbol")
# Column `symbol` joining factors with different levels, coercing to character vector
capm_results <- inner_join(capm_results, rsq_errors, by="symbol")


# rename and keep beta, se(beta), and CI for beta, alpha, rsq, regression SE (specific risk)
capm_results <- capm_results %>% 
  rename(
    beta = estimate.x,
    se_beta = std.error.x,
    beta_low = conf.low.x,
    beta_high = conf.high.x,
    alpha = estimate.y,
    residual_se = sigma) %>% 
  select(symbol, beta, se_beta, beta_low, beta_high, alpha, r.squared, residual_se)


capm_results %>% 
  arrange(desc(beta)) %>% 
  kable()%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```

# Estimating CAPM betas for all of SP500 consituents

## Get ticker symbols off wikipedia

Lets us get a list of all ticker symbols from the SP500

```{r, sp500_tickers_from_wikipedia, warning=FALSE}

sp500_url <- "https://en.wikipedia.org/wiki/List_of_S%26P_500_companies"

#get tables that exist on URL
tables <- sp500_url %>% 
  read_html() %>% 
  html_nodes(css="table")


# parse HTML tables into a dataframe called djia. 
# Use purr::map() to create a list of all tables in URL
sp500 <- map(tables, . %>% 
               html_table(fill=TRUE)%>% 
               clean_names())


# constituents
table1 <- sp500[[1]] %>% # the second table on the page contains the ticker symbols
  mutate(date_first_added = ymd(date_first_added),
         
         # if a stock is listed on NYSE, its symbol is, e.g., NYSE: MMM
         # We will get prices from yahoo finance which requires just the ticker
         
         # if symbol contains "NYSE*", the * being a wildcard
         # then we jsut drop the first 6 characters in that string
         ticker = ifelse(str_detect(symbol, "NYSE*"),
                          str_sub(symbol,7,11),
                          symbol)
         )
# we need a vector of strings with just the tickers
tickers <- table1 %>% 
  select(symbol) %>% 
  pull()  # pull() gets them as a sting of characters
```

## Download prie data using `tidyquant`

```{r get_price_data_sp500, message=FALSE, warning=FALSE, cache=TRUE}
# Notice the cache=TRUE argument in the chunk options. Because getting data is time consuming, 
# cache=TRUE means that once it downloads data, the chunk will not run again next time you knit your Rmd

sp500_stocks <- tickers %>% 
  tq_get(get  = "stock.prices",
         from = starting_date # go back 5 years from today; otherwise, give starting date as "2017-01-01"
         ) %>%
  group_by(symbol)

glimpse(sp500_stocks) # examine the structure of the resulting data frame
```
## Calculate monthly returns

```{r sp500_CAPM, warning=FALSE}
#calculate monthly  returns
my_big_stocks_returns_monthly <- sp500_stocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "monthly_return",
               cols = c(nested.col)) 

#calculate SPY monthly  returns
spy_returns_monthly <- spy %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "SPY_return",
               cols = c(nested.col))

sp500_data <- left_join(my_big_stocks_returns_monthly, spy_returns_monthly, by="date")

too_many_models <- sp500_data %>% 
  # create a list containing a separate dataframe for each stock
  group_by(symbol) %>% 
  nest()  %>% 
  
  # Run a simple regression model for every country in the dataframe
  mutate(simple_capm_model = data %>% 
           map(~lm(monthly_return ~  SPY_return, data = .))) %>% 
  
  # extract regression coefficients and model details with broom::tidy()
  mutate(coefs = simple_capm_model %>% 
           map(~ tidy(., conf.int = TRUE)),
         details = simple_capm_model %>% 
           map(glance)) %>% 
  ungroup()


# pull intercepts, or alphas 
intercepts <- 
  too_many_models %>% 
  unnest(coefs) %>% 
  filter(term == "(Intercept)") %>% 
  arrange(estimate) %>% 
  mutate(symbol = fct_inorder(symbol)) %>% 
  select(symbol, estimate, std.error, conf.low, conf.high)

# let us look at the intercepts, or alphas
# intercepts %>% 
#   rename(alpha = estimate) %>% 
#   kable()%>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


# pull slopes, or betas 
slopes <- too_many_models %>% 
  unnest(coefs) %>% 
  filter(term == "SPY_return") %>% 
  arrange(estimate) %>% 
  mutate(symbol = fct_inorder(symbol)) %>% 
  select(symbol, estimate, std.error, conf.low, conf.high)

# let us look at the betas, or measure of market risk
# slopes %>% 
#   rename(beta = estimate) %>% 
#   arrange(desc(beta)) %>% 
#   kable()%>%
#   kable_styling(bootstrap_options = c("striped", "hover", "condensed"))


# pull R^2 and regression SE, or specific risk 
rsq_errors <- too_many_models %>% 
  unnest(details) %>% 
  select(symbol, r.squared, sigma) %>% 
  arrange(desc(r.squared)) 

# let us join everything together
capm_results <- inner_join(slopes, intercepts, by="symbol")
# Column `symbol` joining factors with different levels, coercing to character vector
capm_results <- inner_join(capm_results, rsq_errors, by="symbol")

capm_results <- inner_join(capm_results, table1, by="symbol")


# rename and keep beta, se(beta), and CI for beta, alpha, rsq, regression SE (specific risk)
# as well as name, sector, and industry
capm_results <- capm_results %>% 
  rename(
    beta = estimate.x,
    se_beta = std.error.x,
    beta_low = conf.low.x,
    beta_high = conf.high.x,
    alpha = estimate.y,
    residual_se = sigma,
    sector = gics_sector) %>% 
  select(symbol, security, beta, se_beta, beta_low, beta_high, alpha, r.squared, residual_se, sector, gics_sub_industry) %>% 
  filter (beta < 5) # get rid of extreme betas


glimpse(capm_results)

# How many companies per sector in the SP500
capm_results %>% 
   count(sector, sort=TRUE)

# How many companies per industry in the SP500
capm_results %>% 
   count(gics_sub_industry, sort=TRUE) 


# plot betas
ggplot(data = capm_results, aes(x = beta, fill=sector))+
  geom_density(alpha = 0.3)+
  theme_bw(8)+
  labs(title = 'Estimating betas of stocks',
        subtitle = paste0(zoo::as.yearmon(starting_date), " to ", zoo::as.yearmon(ending_date)),
       x = "") +
  facet_wrap(~sector)+
  theme(legend.position = "none")+
  NULL


ggplot(data = capm_results, aes(x = sector, y=beta, fill=sector))+
  geom_violin(alpha = 0.4)+
  theme_minimal(8)+
  labs(title = 'Estimating betas of stocks',
       subtitle = paste0(zoo::as.yearmon(starting_date), " to ", zoo::as.yearmon(ending_date)),
       x = "") +
  coord_flip()+
  theme(legend.position = "none")+
  geom_hline(yintercept = 1)+
  NULL


ggplot(capm_results, aes(x = beta, y = sector, fill=sector)) +
  geom_density_ridges(alpha= 0.6) + 
  theme_bw() + 
  labs(
    x = "Beta estimate", 
    y = "",
    title = "Estimating SP500 betas",
    subtitle = paste0(zoo::as.yearmon(starting_date), " to ", zoo::as.yearmon(ending_date)),
  ) + 
  theme(
    axis.title = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    legend.position = "none"
  )+
  geom_vline(xintercept = 1, colour = "#001e62")+
  NULL


capm_results %>% 
  arrange(desc(beta)) %>% 
  kable()%>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

# save file with stock data
write_csv(sp500_stocks, "sp500_stocks.csv")
```
